---
title: "HW2"
author: "Gunnar"
date: "10/3/2021"
output: 
  github_document:
  collapse: TRUE,
  message: FALSE,
  warning: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading libraries
```{r, message=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
```

**Problem 1**
Part 1. 
Loading Mr. Trash Wheel dataset
filtering out row totals
cleaning variable names and removing unnecssary variables
rounding sport balls to the nearest integer
```{r, message=FALSE}
library(readxl)
df <- read_excel("working data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel")

df <- df %>%
  filter(!endsWith(Month, "Total")) %>%
  clean_names() %>%
  select(-x15, -x16, -x17) %>%
  mutate(sports_balls = round(sports_balls))

df %>%
  nrow()

df %>%
  pull(sports_balls) %>%
  head()


df %>%
  group_by(year) %>%
  summarize(mean(chip_bags))

```


Part 2. 

Importing 2018 and 2019 data sets.
Remove first row of data.
Add year variable to each data frame.
Omit rows with missing precipitation data.
Omit total row for each df.
```{r, message=FALSE}
df_p2019 <- read_excel("working data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", skip = 1)

df_p2019 <- df_p2019 %>%
  clean_names %>%
  mutate(year = 2019) %>%
  filter(!is.na(total)) %>%
  filter(!is.na(month))

df_p2018 <- read_excel("working data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1)  

df_p2018 <- df_p2018 %>%
  clean_names %>%
  mutate(year = 2018) %>%
  filter(!is.na(total)) %>%
  filter(!is.na(month))

df_p_combined <- bind_rows(df_p2018, df_p2019) %>%
  mutate(month = month.name[month])


df_p_combined %>%
  group_by(year) %>%
  summarise(mean(total))

nrow(df_p_combined)

```

Part 3. 

For the Mr. Trash Wheels data set, there were 433 observations spanning between 2014 and 2021. In 2015, there was a mean of 1849 chip bags found in a dumpster while in 2019 there was a mean of 1025. 

For the combined precipitation data set, there were 18 observations throughout 2018 and 2019. In 2018, the mean amount of precipitation was 5.86 inches while the mean amount of precipitation was 2.78 inches in 2019.


**Problem 2**
Loading pols-month.csv, unemployment.csv, and snp.csv
```{r, message=FALSE}
pols_df <- read_csv("working data/fivethirtyeight_datasets/pols-month.csv")

unemploy_df <- read_csv("working data/fivethirtyeight_datasets/unemployment.csv")

snp_df <- read_csv("working data/fivethirtyeight_datasets/snp.csv")
```

Cleaning pols_month dataframe
```{r, message=FALSE}
pols_df <- pols_df %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = as.numeric(month)) %>%
  mutate(month = month.name[month]) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(prez = ifelse(prez_gop == 1, "gop", ifelse(prez_dem == 1, "dem", NA))) %>%
  select(-prez_gop, -prez_dem, -day) %>%
  relocate(year, month)


names(pols_df)
summary(pols_df)
head(pols_df$month)
nrow(pols_df)
```

Cleaning snp dataframe
```{r, message=FALSE}
snp_df <- snp_df %>%
  separate(date, into = c("year", "month", "day"), sep = "/") %>%
  mutate(month = as.numeric(month)) %>% 
  mutate(year = as.numeric(year)) %>%
  mutate(month = month.name[month]) %>%
  mutate(year = year + 2000) %>%
  select(-day) %>%
  relocate(year, month)

names(snp_df)
summary(snp_df)
head(snp_df$year)
nrow(snp_df)
```

Cleaning unemployment dataframe
```{r, message=FALSE}
unemploy_df <- unemploy_df %>%
  clean_names() %>%
  pivot_longer(cols = jan:dec, names_to = "month", values_to = "unemployment") %>%
  mutate(month = factor(month, labels = month.name)) %>%
  select(year, month, unemployment)

names(unemploy_df)
summary(unemploy_df)
nrow(unemploy_df)
```

Merging the data sets

```{r, message=FALSE}
pols <- left_join(snp_df, pols_df, by = c("year", "month"))
pols <- left_join(pols, unemploy_df)
head(pols)

pols %>%
  group_by(prez) %>%
  summarise(mean(unemployment))
```

*Discussion of the data set*

After cleaning the data sets and before merging them, the pols data set contained 822 observations over 9 variables: year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem, and prez. The snp data set, meanwhile, contained 787 observations over 3 variables: year, month, and close. Lastly, the unemployment data set has 816 observations: year, month, and unemployment level.

In the combined data set, there are 787 observations over 11 variables. The time range for the combined data set is between 2001 and 2012. An interesting observation from the data set is that the average unemployment rate under democratic presidents was 9.00 while it was 5.28 under republican presidents. 

**Problem 3**

Loading the baby name data set
```{r}
baby_df <- read_csv("working data/Popular_Baby_Names.csv")
```

Cleaning the baby name data set
First, clean names
Second, make all values lower case
Third, merge name versions
Fourth, remove all duplicate rows (based on repeat of name, year, and gender)
```{r, message=FALSE}
baby_df <- baby_df %>%
  clean_names() %>%
  mutate(gender = tolower(gender), ethnicity = tolower(ethnicity), 
        childs_first_name = tolower(childs_first_name)) %>%
  mutate(ethnicity = ifelse(ethnicity == "asian and paci" | ethnicity == "asian and pacific islander", "asian and pacific islander", ifelse(ethnicity == "white non hispanic" | ethnicity == "white non hisp", "white non hispanic", ifelse(ethnicity == "black non hispanic" | ethnicity == "black non hisp", "black non hispanic", ethnicity)))) %>%
  distinct(childs_first_name, gender, year_of_birth, ethnicity, .keep_all = TRUE)

baby_df %>%
  filter(childs_first_name == "olivia")
```

Creating table of frequency of name Olivia  over time
Filter to only girls named Olivia
group by birth year
pivot wider so that years are columns
```{r, message=FALSE}
olivia_table <- baby_df %>%
  filter(childs_first_name == "olivia", gender == "female") %>%
  select(-count) %>%
  group_by(year_of_birth) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank) %>%
  select(ethnicity, everything(), -gender, -childs_first_name) %>%
  knitr::kable(caption = "Popularity Ranking of the Name Olivia among New Born Girls over Time"
  )

olivia_table
```

Creating table of top boys name over time
```{r, message=FALSE}
boys_table <- baby_df %>%
  filter(gender == "male", rank == 1) %>%
  select(-count) %>%
  group_by(year_of_birth) %>%
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name) %>%
  select(ethnicity, everything(), -gender, -rank) %>%
  knitr::kable(caption = "Most Popular Boy's Name by Year and Ethnicity"
  )

boys_table
```

Scatter plot of male, white non-hispanic children born in 2016 of name count vs name rank
```{r, message=FALSE}
boys_name_p <- baby_df %>%
  filter(ethnicity == "white non hispanic", year_of_birth == 2016) %>%
  ggplot(aes(x = count, y = rank)) +
  geom_point(position = "jitter", alpha = 0.5) +
  labs(title = "Non Hispanic White Boy's Name Frequency vs Rank in 2016") +
       xlab("Name Count") +
       ylab("Name Rank")

boys_name_p
```

